---
title: "Final Report"
author: "Aoi Ogawa, Ahlam El Mernissi, Christina Lyu"
date: "12/12/2018"
output:
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(ggplot2)
library(mosaic)
library(leaps)
library(car)
```

### Abstract:
Theabstractshould be a 4-8 sentence paragraph that clearly (but concisely) describes your main research questions, the type of data you have, and your primary findings. 
  Our research question is "What are some common characteristics for billioniares?". Our assumption is that most billionaires' wealth is inherited and they are mainly male. The data we use has a quantative response variable and several categorical predictors. The test statistics show that billionaires with business in the fields including Communications, Technology tend to have a significant higher net worth than others. It has also been found that the two variables Year and Rank explain the most variance in the net worth of billionaires. 

### Introduction: 

	Everyone has a billionaire dream. Having money means living in mansions, flying private planes, attending parties and meeting celebrities. It is very unlikely to know a billionaire personally, but we can learn about their life stories from books and interviews. They might not be true, but they gives us some characters to look up to in case someday we might also become billionaires. We assume that billionaires share similar characteristics, which are not possessed by the general public and are essential to the outcome of becoming billionaires. Therefore, we want to investigate the shared characteristics of billionaires.
	
### Data:

  The data we found is called “The Billionaire Characteristics Database” from the website OpenDataSoft. It examines the billionaires listed on Forbes for 20 years and collect information about them and their companies according to the magazine Rich People Poor Countries.
  
  The observations in the dataset are each billionaire and the columns contain information about themselves and their companies, including age, gender, the industry their business mainly dwell in and many more. To explore the relationship between billionaires and the common characteristics, we use their net worth Net.Worth.Billion as the response variable and the others as predictors. One predictor is Rank, which is the rank of the billionaires in Forbes. As they are ranked based on their net worth, Rank is clearly strongly related to Net.Worth.Billion, but it is not a very informative predictor to interpret the common characteristics of the billionaires. 
  
  The cleansing process of the data includes excluding rows with N.A and 0 values, creating additional binary predictors based on existing ones and fit models according to the new variables. 
  
  
### Methods:

```{r, message=FALSE, warning=FALSE}
# read the csv file 
billionaire <- read.csv("the-billionaire-characteristics-database.csv")

#tidy up the dataset 
tidy_billionaire <- subset(billionaire, select = -c(3, 4, 5, 10, 15, 17, 21, 23, 24, 25,26))

# change the blank cell into a NA
tidy_billionaire[tidy_billionaire == ""] <- NA

without_na <- na.omit(tidy_billionaire)

# excluded NA for gender
na_gender <- subset(tidy_billionaire, !is.na(Gender))

# excluded NA for selfmade and exclude NA for gender
gender_n_Selfmade <- subset(na_gender, !is.na(Selfmade))

# change NA in Political.Connections into 0
gender_n_Selfmade$Political.Connection[is.na(gender_n_Selfmade$Political.Connection)] <- 0

# create a age group (youth, young_adult, adult, senior)
age_billionare <- gender_n_Selfmade %>% 
  mutate(age_group = ifelse((Age >= -42 & Age <= 18 & Age == 0), 'youth', ifelse((Age > 18 & Age <= 35), 'young_adult', ifelse((Age > 35 & Age <= 55), 'adult', ifelse((Age > 55), 'senior', 'other')))))

# removing age_group NA valuables
age_group_na <- subset(age_billionare, !is.na(age_group))

#removing founding year NA variables
founding_year_na <- subset(age_group_na, !is.na(Founding.Date))

#age of company 
final_data <- founding_year_na %>%
  mutate(company_age = 2016 - Founding.Date)


# remove all the na 
na_billionare <- na.omit(tidy_billionaire)
# there are only 61 observations when removed all the na from the dataset we removed some columns => not a resonable thing to do
```

```{r}
# correlation between variables

# pairs(gender_n_Selfmade)
cor(gender_n_Selfmade[,unlist(lapply(gender_n_Selfmade, is.numeric))])
```


```{r,warning=FALSE, message=FALSE}
# color code by gender and shape by Selfmade or not 
ggplot(gender_n_Selfmade, aes(x = Age, y = Net.Worth.Billion, color = Gender, shape = Selfmade)) +
  geom_point()


# model of Age, Gender and Selfmade predicting the Net.Worth.Billion
mod_age_gen_self <- lm(Net.Worth.Billion ~ Age + Gender + Selfmade, gender_n_Selfmade)

summary(mod_age_gen_self)
```

```{r,warning=FALSE, message=FALSE}
gender_n_Selfmade2 <- gender_n_Selfmade[gender_n_Selfmade$Industry!="0.0", ]

# color code by Industry
ggplot(gender_n_Selfmade2, aes(x = Founding.Date, y = Net.Worth.Billion, color = Industry)) + 
  geom_point()

ggplot(gender_n_Selfmade2, aes(x = Industry, y = Net.Worth.Billion)) + geom_point()


# model of Founding date and Industry predicting the Net.Worth.Billion
mod_date_ind <- lm(Net.Worth.Billion ~ Founding.Date + Industry, gender_n_Selfmade2)

summary(mod_date_ind)

```

```{r,warning=FALSE, message=FALSE}

# color code by region
ggplot(gender_n_Selfmade, aes(x = Founding.Date, y = Net.Worth.Billion, color = Region)) + 
  geom_point()


# model of Founding date and region predicting the Net.Worth.Billion
mod_date_reg <- lm(Net.Worth.Billion ~ Founding.Date + Region, gender_n_Selfmade)

summary(mod_date_reg)

```

```{r,warning=FALSE, message=FALSE}

#color code by political connection
ggplot(gender_n_Selfmade, aes(x = Founding.Date, y = Net.Worth.Billion, color = factor(Political.Connection))) + 
  geom_point()


# model of Founding date and Political.Connection predicting the Net.Worth.Billion



mod_date_poli <- lm(Net.Worth.Billion ~ Founding.Date + factor(Political.Connection), gender_n_Selfmade)

summary(mod_date_poli)

```

```{r,warning=FALSE, message=FALSE}

# color code by the type of company 
ggplot(gender_n_Selfmade, aes(x = Age, y = Net.Worth.Billion, color = Company.type)) +
  geom_point()


# model of Founding date and Industry predicting the Net.Worth.Billion
mod_age_comp <- lm(Net.Worth.Billion ~ Age + Company.type, gender_n_Selfmade)

summary(mod_age_comp)
```

```{r,warning=FALSE, message=FALSE}

# Separate by each company type  
ggplot(gender_n_Selfmade, aes(x = Age, y = Net.Worth.Billion, color = Gender)) +
  geom_point() + 
  facet_grid(. ~Company.type)

```


```{r}
gender_n_Selfmade[complete.cases(gender_n_Selfmade), ]
```

```{r}

a <- aov(Net.Worth.Billion ~ Type.of.Wealth, data = gender_n_Selfmade)
summary(a)
# color code by the type of company 
ggplot(gender_n_Selfmade, aes(x = Founding.Date, y = Net.Worth.Billion, color = Type.of.Wealth)) +
  geom_point()
lm <- lm(log10(Net.Worth.Billion) ~ Type.of.Wealth, data = gender_n_Selfmade)
summary(lm)
plot(lm)
```

```{r}

names(final_data) <-c("Year", "Rank", "Net_Worth_Billion", "Age", "Industry_Aggregates", "Selfmade", "Type_of_Wealth", "Industry", "Relationship_to_Company", "Gender", "Sector", "Company_type", "Founding_Date", "Political_Connection", "Region", "Characteristics", "age_group", "company_age")
```

```{r}
# construct stepwise selection to get significant variables
stepwise <- regsubsets(Net_Worth_Billion ~ Year + Rank + Industry_Aggregates + Selfmade + Type_of_Wealth + Industry + Relationship_to_Company + Gender + Sector + Company_type + Founding_Date + Political_Connection + Region + Characteristics + age_group + company_age, data = final_data, nbest = 1, really.big = TRUE, method = "seqrep")

with(summary(stepwise), data.frame(rsq, adjr2, cp, rss, bic, outmat))
```

```{r}
final_data[final_data$Sector == "Communications", ]
```


```{r}
final_data_1 <- final_data %>%
  mutate(SectorCommunications = ifelse(Sector == "Communications", 1, 0)) %>%
  mutate(SectorFashion = ifelse(Sector == "fashion", 1, 0)) %>%
  mutate(SectorTechnology = ifelse(Sector == "technology", 1, 0)) %>%
  mutate(SectorFinance = ifelse(Sector == "finance", 1, 0)) %>%
  mutate(SectorSoftware = ifelse(Sector == "software", 1, 0)) %>%
  mutate(Company_typeacquired = ifelse(Company_type == "acquired", 1, 0)) %>%
  mutate(SectorOilRefining = ifelse(Sector == "oil refining", 1, 0))
```


```{r}
m <- lm(Net_Worth_Billion ~ Year + Rank + SectorCommunications + SectorFashion + SectorFinance + SectorSoftware + SectorTechnology + Company_typeacquired + SectorOilRefining, data = final_data_1)
summary(m)
```

```{r}
mod_age <- lm( Net_Worth_Billion ~ Age, final_data_1) 
plot(mod_age)
plot(mod_age, which = 4)
plot(m, which = 4)
```


```{r}

#billionare age dividing by region
ggplot(final_data_1, aes(x = Age, y = Net_Worth_Billion)) + geom_point() + facet_grid(. ~Region) + geom_smooth(method = 'lm', se = FALSE)

```

```{r}
#median age for the each region
East_Asia <- subset(final_data, Region == "East Asia")
median(East_Asia$Age)

Europe <- subset(final_data, Region == "Europe")
median(Europe$Age)

Latin_America <- subset(final_data, Region == "Latin America")
median(Latin_America$Age)

Middle_East.North_Africa <- subset(final_data, Region == "Middle East/North Africa")
median(Middle_East.North_Africa$Age)

North_America <-subset(final_data, Region == "North America")
median(North_America$Age)

South_Asia <- subset(final_data, Region == "South Asia")
median(South_Asia$Age)

Subsaharan_Africa <-subset(final_data, Region == "Sub-Saharan Africa")
median(Subsaharan_Africa$Age)

```




```{r}
#transformed Net_Worth_Billion
m2 <- lm(log(Net_Worth_Billion) ~ Year + Rank + SectorCommunications + SectorFashion + SectorFinance + SectorSoftware + SectorTechnology + Company_typeacquired, data = final_data_1)
summary(m2)
plot(m2)
```





```{r}
require(mosaic)
require(Stat2Data)
require(dplyr)
set.seed(12345)
```

```{r}

# omit NA values
NA_final_data <- na.omit(final_data_1)

# coefficients of the original data 
cor(Net_Worth_Billion ~ Year + Rank + SectorCommunications + SectorFashion + SectorFinance + SectorSoftware + SectorTechnology + Company_typeacquired, data = NA_final_data)

# point estimate??
cor(Net_Worth_Billion ~ shuffle(Year + Rank + SectorCommunications + SectorFashion + SectorFinance + SectorSoftware + SectorTechnology + Company_typeacquired), data = NA_final_data)

#randomization
rand.test1 <- do(5000) * cor(Net_Worth_Billion ~ shuffle(Year + Rank + SectorCommunications + SectorFashion + SectorFinance + SectorSoftware + SectorTechnology + Company_typeacquired), data = NA_final_data)

densityplot(~cor, data=rand.test1)

summary(rand.test1)


```



```{r}
#bootstrap
#bootstrap <- do(5000) * coef(lm(Net_Worth_Billion ~ Year + Rank + SectorCommunications + SectorFashion + SectorFinance + SectorSoftware + SectorTechnology + Company_typeacquired, data = final_data))
#densityplot( ~ Intercept, data = bootstrap)
#summary(bootstrap)
#bootstrap %>% summarize(ds = sd(final_data$Net_Worth_Billion))
```


$$

\ Net\_Worth\_Billion = \beta_0 + \beta_1 \cdot \ Year \beta_2 \cdot \ Rank + \beta_3 \cdot	\ SectorCommunications + \beta_4 \cdot \SectorFashion + \beta_5 \cdot \SectorFinance + \beta_6 \cdot \SectorSoftware \beta_7 \cdot	\SectorTechnology \beta_8 \cdot \Company_typeacquired \beta_8 \cdot \ SectorOilRefining

$$
--------------------------------------------------------------------------------------------------------------
creating another model

```{r}

# without rank and year
new_stepwise <- regsubsets(Net_Worth_Billion ~ Industry_Aggregates + Selfmade + Type_of_Wealth + Industry + Relationship_to_Company + Gender + Sector + Company_type + Founding_Date + Political_Connection + Region + Characteristics + age_group + company_age, data = final_data, nbest = 1, nvmax = 24, really.big = TRUE, method = "seqrep")

with(summary(new_stepwise), data.frame(rsq, adjr2, cp, rss, bic, outmat))
```


```{r}
#create new data frame for doing anova
new_final_data <- final_data %>%
  mutate(Communications = ifelse(Sector == "Communications", 1, 0)) %>%
  mutate(Fashion = ifelse(Sector == "fashion", 1, 0)) %>%
  mutate(Technology = ifelse(Sector == "technology", 1, 0)) %>%
  mutate(Finance = ifelse(Sector == "finance", 1, 0)) %>%
  mutate(Software = ifelse(Sector == "software", 1, 0)) %>%
  mutate(Acquired = ifelse(Company_type == "acquired", 1, 0)) %>%
  mutate(Oil_refining = ifelse(Sector == "oil refining", 1, 0)) %>%
  mutate(Chairman = ifelse(Relationship_to_Company == "chairman", 1, 0)) %>% 
  mutate(Construction_real_estate_banking = ifelse(Sector == "construction/real estate/banking", 1, 0)) %>%
  mutate(age_senior = ifelse(age_group == "senior", 1, 0))

new_final_data %>% na.omit(new_final_data)

# each model which the new_stepwise chose 
mod_1 <-lm(Net_Worth_Billion ~ Software, new_final_data)

mod_2 <-lm(Net_Worth_Billion ~ Communications + Software, new_final_data)

mod_3 <-lm(Net_Worth_Billion ~ Communications + Fashion + Software, new_final_data)

mod_4 <-lm(Net_Worth_Billion ~ Communications + Fashion + Software + Acquired, new_final_data)

mod_5 <-lm(Net_Worth_Billion ~ Communications + Fashion + Software + Technology + Acquired, new_final_data)

mod_6 <-lm(Net_Worth_Billion ~ Communications + Fashion + Finance + Software + Technology + Acquired, new_final_data)

mod_7 <-lm(Net_Worth_Billion ~ Communications + Fashion + Finance + Oil_refining + Software + Technology + Acquired, new_final_data)

mod_8 <-lm(Net_Worth_Billion ~ Chairman + Communications + Fashion + Finance + Oil_refining + Software + Technology + Acquired, new_final_data)

mod_9 <-lm(Net_Worth_Billion ~ Chairman + Communications + Construction_real_estate_banking + Fashion + Finance + Oil_refining + Software + Technology + Acquired, new_final_data)

mod_10 <- lm(Net_Worth_Billion ~ Chairman + Communications + Construction_real_estate_banking + Fashion + Finance + Oil_refining + Software + Technology + Acquired + age_senior, new_final_data)

anova(mod_1, mod_2, mod_3)
anova(mod_5, mod_6, mod_7, mod_8, mod_9, mod_10)

summary(mod_4)

summary(mod_1)

#significant 
summary(mod_2)

summary(mod_3)
summary(mod_4)

summary(mod_5)

summary(mod_6)

summary(mod_7)

#significant the best I guess from looking at the summary and anova 
summary(mod_9) 

#significant
summary(mod_10)
```

```{r}
vif(mod_9)
# VIF is low so no multicolinearity 
```

-----------------------------------------------------------------------------------------------------------------------
looking at bilinares on 2014 (present, recent?)

```{r}
# select just 2014 data
data_2014 <- filter(final_data, Year == "2014")

# not including year and rank
stepwise_2016 <- regsubsets(Net_Worth_Billion ~ Industry_Aggregates + Selfmade + Type_of_Wealth + Industry + Relationship_to_Company + Gender + Sector + Company_type + Founding_Date + Political_Connection + Region + Characteristics + age_group + company_age, data = final_data, nbest = 1, really.big = TRUE, method = "seqrep")

with(summary(stepwise_2016), data.frame(rsq, adjr2, cp, rss, bic, outmat))

```

```{r}
new_2014_data <- data_2014 %>%
  mutate(Communications = ifelse(Sector == "Communications", 1, 0)) %>%
  mutate(Fashion = ifelse(Sector == "fashion", 1, 0)) %>%
  mutate(Technology = ifelse(Sector == "technology", 1, 0)) %>%
  mutate(Finance = ifelse(Sector == "finance", 1, 0)) %>%
  mutate(Software = ifelse(Sector == "software", 1, 0)) %>%
  mutate(Acquired = ifelse(Company_type == "acquired", 1, 0)) %>%
  mutate(Oil_refining = ifelse(Sector == "oil refining", 1, 0)) %>%
  mutate(Chairman = ifelse(Relationship_to_Company == "chairman", 1, 0)) %>% 
  mutate(Construction_real_estate_banking = ifelse(Sector == "construction/real estate/banking", 1, 0)) 

new_2014_data %>% na.exclude(new_2014_data)

present_1 <-lm( Net_Worth_Billion ~ Software, new_2014_data)
  
present_2 <-lm( Net_Worth_Billion ~ Communications + Software, new_2014_data)
  
present_3 <-lm( Net_Worth_Billion ~ Communications + Fashion + Software, new_2014_data)
  
present_4 <-lm( Net_Worth_Billion ~ Communications + Fashion + Software + Acquired, new_2014_data)

present_5 <-lm( Net_Worth_Billion ~ Communications + Fashion + Software + Technology + Acquired, new_2014_data)

present_6 <-lm(Net_Worth_Billion ~ Communications + Fashion + Finance + Software + Technology + Acquired, new_2014_data)

present_7 <-lm(Net_Worth_Billion ~ Communications + Fashion + Finance + Oil_refining + Software + Technology + Acquired, new_2014_data)

present_8 <-lm(Net_Worth_Billion ~ Chairman + Communications + Fashion + Finance + Oil_refining + Software + Technology + Acquired, new_2014_data)

present_9 <- lm(Net_Worth_Billion ~ Chairman + Communications + Construction_real_estate_banking + Fashion + Finance + Oil_refining + Software + Technology + Acquired, new_2014_data)

anova(present_1, present_2, present_3) 
anova(present_4, present_5, present_6, present_7, present_8, present_9)

summary(present_2)

summary(present_5)

summary(present_9)

# not sure which model is significant 
```

```{r}
# looking at billionares in 2014
# color by sector 
ggplot(new_2014_data, aes(x = age_group, y = Net_Worth_Billion, color = Sector)) + geom_point()

# Looking at region
new_2014_data$age_group <- factor(new_2014_data$age_group, levels = c("young_adult", "adult", "senior"))
ggplot(new_2014_data, aes(x = age_group, y = Net_Worth_Billion, color = Region)) + geom_point()
```

### Statistics:

### Results:

### Conclusions:

### Reference:

1. “The Billionaire Characteristics Database.” OpenDataSoft -, `data.opendatasoft.com/explore/dataset/the-billionaire-characteristics-database@public-us/table/`.

